---
title: "newGroup"
output: pdf_document
---
---
title: "data explore"
author: "Wanying Ma"
date: "1/7/2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{bm}
- \usepackage{amsmath}
- \usepackage{booktabs}
- \usepackage{pgfplots}
- \usepackage{pgfplotstable}
- \usepackage{multirow}
- \usepackage{bm}
- \usepackage{amsmath}
- \usepackage{booktabs}
- \usepackage{pgfplots}
- \usepackage{pgfplotstable}
- \usepackage{multirow}
- \usepackage{bm}
- \newcommand{\Real}{\mathbb{R}}
- \newcommand{\dom}{{\bf dom}\,}
- \newcommand{\Tra}{^{\sf T}} % Transpose
- \newcommand{\Inv}{^{-1}} % Inverse
- \def\vec{\mathop{\rm vec}\nolimits}
- \def\sweep{\mathop{\rm sweep}\nolimits}
- \newcommand{\diag}{\mathop{\rm diag}\nolimits}
- \newcommand{\tr}{\operatorname{tr}} % Trace
- \newcommand{\epi}{\operatorname{epi}} % epigraph
- \newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
- \newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
- \newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
- \newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
- \newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
- \newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element
- \newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
- \newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
- \newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
- \newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
- \newcommand{\Mcheck}[1]{{\bm{\check \mathbf{\MakeUppercase{#1}}}}} % matrix
- \newcommand{\Mbar}[1]{{\bm{\bar \mathbf{\MakeUppercase{#1}}}}} % matrix
- \newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
---

## In this project, we intend to use the DTI scanning result on 27 tracts and 4 measurements for 195 players to cluster/classify 3 groups of players. The 3 groups stands for players who had head injury in the game, players at the same spot in the body contact game but without having head injury, and player at the same spot in the noncontact game without having head injury.

## Access to data files

```{r, echo=FALSE, warning=FALSE}
library(dplyr)
library(refund)
# there are 195 players
files <- list.files(path="/Users/mawanying/moon\ Dropbox/Wanying\ Ma/Luo_Jarek_YC/csv_voxelwise", pattern="*.csv", full.names=FALSE, recursive=FALSE)
#files <- list.files(path="C:/Users/wma9/Dropbox/Luo_Jarek_YC/csv_voxelwise", pattern="*.csv", full.names=FALSE, recursive=FALSE)

subname <- gsub(pattern=".csv", replacement="", files)
# there are 27 tracts 
tractnames = c("ar_l",  "ar_r",  "atr_l", "atr_r", "cgc_l", "cgc_r", "cgh_l", "cgh_r", "cst_l", "cst_r","fma",   
               "fmi", "ifo_l", "ifo_r", "ilf_l", "ilf_r", "mcp",  
               "ml_l",  "ml_r",  "ptr_l", "ptr_r", "slf_l", "slf_r", "str_l" ,"str_r", "unc_l", "unc_r")
# 200 quantiles we want to use
den.quantiles = seq(0.25, 99.75, length=200)
measure = c("FA","MD","Da","Dr")
para <- expand.grid(t=tractnames, m=measure)


load("//wolftech.ad.ncsu.edu/cos/stat/Redirect/wma9/Desktop/fromdropbox/football project/alldata_qdtg3.Rdata") #use new 73 players to do normalization
load("//wolftech.ad.ncsu.edu/cos/stat/Redirect/wma9/Desktop/fromdropbox/football project/summary_data.Rdata")
load("//wolftech.ad.ncsu.edu/cos/stat/Redirect/wma9/Desktop/fromdropbox/football project/alldata_qdt.Rdata") #dim(alldata_qdt): 21052x203
load("//wolftech.ad.ncsu.edu/cos/stat/Redirect/wma9/Desktop/fromdropbox/football project/alldata_q.Rdata") #dim(alldata_qdt): 21052x203
load("//wolftech.ad.ncsu.edu/cos/stat/Redirect/wma9/Desktop/fromdropbox/football project/alldata_qdtg.Rdata") #dim(alldata_qdt): 21052x203


load("/Users/mawanying/moon\ Dropbox/Wanying\ Ma/fromdropbox/alldata_q.Rdata") #dim(alldata_qdt): 7876x203
load("/Users/mawanying/moon\ Dropbox/Wanying\ Ma/fromdropbox/football project/summary_data.Rdata")
load("/Users/mawanying/moon\ Dropbox/Wanying\ Ma/fromdropbox/football project/alldata_qdt.Rdata") 
load("/Users/mawanying/moon\ Dropbox/Wanying\ Ma/fromdropbox/alldata_qdtg5.Rdata") 

```

## Data preprocessing: 1. Add group index(merge with player information); 2. Keep only football players, and delete duplicates between 3 groups.
```{r , echo=FALSE, warning=FALSE}
#############################################################################
######################### players information table #########################
######################### create group index ################################
######################### delete duplicates #################################
#############################################################################
player.info <- read.csv("/Users/mawanying/moon\ Dropbox/Wanying\ Ma/Luo_Jarek_YC/MRI_Injury_Description_03OCT18_All_Concussions_ForLuo.csv") #player.info is 136x9
files <- list.files(path="/Users/mawanying/moon\ Dropbox/Wanying\ Ma/Luo_Jarek_YC/csv_voxelwise", pattern="*.csv", full.names=FALSE, recursive=FALSE)

#player.info <- read.csv("C:/Users/wma9/Dropbox/Luo_Jarek_YC/MRI_Injury_Description_03OCT18_All_Concussions_ForLuo.csv") #player.info is 136x9
#files <- list.files(path="C:/Users/wma9/Dropbox/Luo_Jarek_YC/csv_voxelwise", pattern="*.csv", full.names=FALSE, recursive=FALSE)
subname <- gsub(pattern=".csv", replacement="", files)
player.info <- player.info %>% filter(INJURYSPORTVARSITY=='Football') #88x9

group1 <- as.character(unique(player.info$SUBJECTSTUDYNUM)) #73 # there is no empty in group1
group1 <- group1[group1!=''] #73
group2 <- as.character(unique(player.info$CONTACTSPORTQGID)) #60
group2 <- group2[group2!=''] #59
group3 <- as.character(unique(player.info$NONCONTACTSPORTQGID)) #59
group3 <- group3[group3!=''] #58


d1 <- group1 %in% group2 #8 there are 8 football players that occurs in both group 1 and group 2
sum(d1)
d2 <- group1 %in% group3 #0
sum(d2)
d3 <- group2 %in% group1 #8 there are 8 football players that occurs in both group 1 and group 2
sum(d3)
d4 <- group2 %in% group3 #0
sum(d4)
d5 <- group3 %in% group1 #0
sum(d5)
d6 <- group3 %in% group2 #0
sum(d6)

group1[which(d1==TRUE)] #c("UCLA-FB-1098", "UCLA-FB-1102", "UCLA-FB-1670", "UNC-FB-1093",  "WISC-FB-1934", "WISC-FB-1937", "WISC-FB-1938", "WISC-FB-1939")
group2[which(d3==TRUE)] #c("UCLA-FB-1098", "UCLA-FB-1102", "UCLA-FB-1670", "UNC-FB-1093",  "WISC-FB-1934", "WISC-FB-1937", "WISC-FB-1938", "WISC-FB-1939") all are fb, and 4 of them are in 195


group1 <- group1[which(d1==FALSE)] #65
group2 <- group2[which(d3==FALSE)] #51

sum(group1 %in% subname) #39
sum(group2 %in% subname) #34
sum(group3 %in% subname) #34

group1 <- group1[group1 %in% subname]#39
group2 <- group2[group2 %in% subname]#34
group3 <- group3[group3 %in% subname]#34

#############################################################################
############## construct group index for football players ###################
##############      group1: SUBJECTSTUDYNUM                 ############
##############      group2: CONTACTSPORTQGID                ############
##############      group3: NONCONTACTSPORTQGID             ############
#############################################################################

# after merge, alldata_fb will have 28 players in group 1, 31players in group 2, 29 players in group 3
#############################################################################
addgroupindex <- function(data){
  alldata.player <- as.character(data$player) #21052x1
  alldata.fb.player <- alldata.player %in% c(group1, group2, group3) #indicator vector: 9496x1
  groupindex <- c()
  for(i in alldata.player[alldata.fb.player]){
    if(i %in% group1){
      groupindex <- c(groupindex, 1)
    }
    if(i %in% group2){
      groupindex <- c(groupindex, 2)
    }
    if(i %in% group3){
      groupindex <- c(groupindex, 3)
    }
  }
  fbdata <- data[alldata.fb.player, ]
  colnames(fbdata)[4:203] <- paste0("den", 1:200)
  data.fb <- fbdata %>% mutate(group=groupindex)
  colnames(data.fb)[4:203] <- paste0("den", 1:200)
  return(data.fb)
}

alldata.fb <- addgroupindex(alldata) #11548x204
alldata.fb_qdt <- addgroupindex(alldata_qdt) #11548x204
alldata.fb_q <- addgroupindex(alldata_q) #11548x204
alldata.fb_qdtg <- addgroupindex(alldata_qdtg)
colnames(alldata.fb)[4:203] <- colnames(alldata.fb_qdt)[4:203] <- colnames(alldata.fb_q)[4:203] <- colnames(alldata.fb_qdtg)[4:203] <- paste0("den", 1:200)
#table(alldata.fb$group)
#1    2    3 
#3016 3348 3132 
#length(unique(alldata.fb$player[alldata.fb$group==1])) #28
#length(unique(alldata.fb$player[alldata.fb$group==2])) #31
#length(unique(alldata.fb$player[alldata.fb$group==3])) #29
# there are 88 different players in total for alldata.fb
#alldata_fb will have 28 players in group 1, 31 players in group 2, 29 players in group 3
```

## cluster
### Using equally-spaced density(no transformation), as well as quantile-density transformation of density $-log(f(Q(t)))$
```{r, echo=FALSE, warning=FALSE}
fPCs <- function(data, iter){
  # data: input dataset
  # m: measurement, in {"FA", "MD", "Da", "Dr"}
  # t: tract
  t <- as.character(para[iter, 1])
  m <- as.character(para[iter, 2])
  data.tm <- data %>% filter(tract==t, measure==m, group !=3) #86x204
  den.tm <- select(data.tm, den1:den200)    #86x200
  den.tm <- den.tm - as.vector(rep(1, dim(den.tm)[1])) %*% t(colMeans(den.tm))
  den.fpca <- refund::fpca.face(data.matrix(den.tm), center = TRUE, argvals = (1:200)/200, knots = 7, pve = .99, p = 5, lambda = 0)
  dim(den.fpca$efunctions)[2]
}
fPCss1 <- fPCss2 <- fPCss3 <- fPCss4 <- c()
for (i in 1:108){
  fPCss1 <- c(fPCss1, fPCs(alldata.fb, i))
  fPCss2 <- c(fPCss2, fPCs(alldata.fb_q, i))
  fPCss3 <- c(fPCss3, fPCs(alldata.fb_qdt, i))
  fPCss4 <- c(fPCss4, fPCs(alldata.fb_qdtg, i))
}

#den .95:
#fPCss
# 2  3  4  5 
#26 59 20  3 

#q .95
#fPCss
# 2  3  4 
#15 73 20

#fPCss .95
#qdt
# 4  5  6  7  8 
# 3 19 36 43  7 

#qdtg .95:
#fPCss
# 4  5  6  7  8 
# 9 18 31 39 11

#######################
#den .99:
#fPCss
#3  4  5  6 
#16 58 29  5 

#q .99
#fPCss
# 4  5  6  7 
#29 48 30  1 

#fPCss .99
#qdt
# 7  8  9 10 11 
# 2 18 35 50  3

#qdtg .99:
#fPCss
# 7  8  9 10 11 
# 5 18 37 44  4 

###################### only group 1 and group 2 ################
#only 2 clusters

#den .95:
#fPCss
# 2  3  4  5 
#32 57 16  3  

#q .95
#fPCss
# 2  3  4 
#10 73 25 

#fPCss .95
#qdt
# 4  5  6  7  8 
# 3 22 34 44  5

#qdtg .95:
#fPCss
#  4  5  6  7  8 
# 7 20 32 41  8

#######################
#den .99:
#fPCss
#3  4  5  6 
#18 57 28  5  

#q .99
#fPCss
#3  4  5  6  7 
#1 24 48 34  1

#fPCss .99
#qdt
# 7  8  9 10 11 
# 3 22 38 44  1

#qdtg .99:
#fPCss
# 7  8  9 10 11 
# 4 21 44 38  1 


m.tract.cluster <- function(data, iter){
  # data: input dataset
  # m: measurement, in {"FA", "MD", "Da", "Dr"}
  # t: tract
  t <- as.character(para[iter, 1])
  m <- as.character(para[iter, 2])
  data.tm <- data %>% filter(tract==t, measure==m) #88x204
  group.tm <- data.tm$group
  den.tm <- select(data.tm, den1:den200)    #88x200
  den.tm <- den.tm - as.vector(rep(1, dim(den.tm)[1])) %*% t(colMeans(den.tm))
  den.fpca <- refund::fpca.face(data.matrix(den.tm), center = TRUE, argvals = (1:200)/200, knots = 7, pve = 0.95, p = 5, lambda = 0)
  den.score <- den.fpca$scores/sqrt(dim(den.tm)[2]) #86x5
  den.cluster <- kmeans(den.score, centers = 3)
 
permute <- function(vec, p1, p2, p3){
  for(i in 1:length(vec)){
    if(vec[i]==1){
      vec[i] <- p1
    }else{
      if(vec[i]==2){vec[i] <- p2}else{
        vec[i] <- p3
      }
    }
  }
  return(vec)  
}


  perm1 <- permute(den.cluster$cluster, 1,2,3)
  perm2 <- permute(den.cluster$cluster, 1,3,2)
  perm3 <- permute(den.cluster$cluster, 2,1,3)
  perm4 <- permute(den.cluster$cluster, 2,3,1)
  perm5 <- permute(den.cluster$cluster, 3,1,2)
  perm6 <- permute(den.cluster$cluster, 3,2,1)
  misrate <- min(sum(perm1!=data.tm$group)/length(data.tm$group),
      sum(perm2!=data.tm$group)/length(data.tm$group),
      sum(perm3!=data.tm$group)/length(data.tm$group),
      sum(perm4!=data.tm$group)/length(data.tm$group),
      sum(perm5!=data.tm$group)/length(data.tm$group),
      sum(perm6!=data.tm$group)/length(data.tm$group))
  data.tm.comb = cbind(data.tm, den.score)
  return(list(misrate=misrate,
              data.tm.comb=data.tm.comb,
              den.fpca=den.fpca))
}

m.tract.cluster2 <- function(data, iter){
  # data: input dataset
  # m: measurement, in {"FA", "MD", "Da", "Dr"}
  # t: tract
  t <- as.character(para[iter, 1])
  m <- as.character(para[iter, 2])
  
  data.tm <- data %>% filter(tract==t, measure==m, group != 3) #86x204
  den.tm <- select(data.tm, den1:den200)    #86x200
  #den.tm <- den.tm - as.vector(rep(1, dim(den.tm)[1])) %*% t(colMeans(den.tm))
  den.fpca <- refund::fpca.face(data.matrix(den.tm), center = TRUE, argvals = (1:200)/200, knots = 7, pve = 0.95, p = 5, lambda = 0)
  den.score <- den.fpca$scores/sqrt(dim(den.tm)[2]) #86x5
  mis <- c()
  for(j in 1:20){
    den.cluster <- kmeans(den.score, centers = 2, iter.max = 100)
    permute <- function(vec, p1, p2){
    for(i in 1:length(vec)){
      if(vec[i]==1){vec[i] <- p1}else{vec[i] <- p2}
    }
    return(vec)
  }
  perm1 <- permute(den.cluster$cluster, 1,2)
  perm2 <- permute(den.cluster$cluster, 2,1)
  misrate <- min(sum(perm1!=data.tm$group)/length(data.tm$group),
                 sum(perm2!=data.tm$group)/length(data.tm$group))
  mis <- c(mis, misrate)
  }
  
  data.tm.comb = cbind(data.tm, den.score)
  return(list(misrate=min(mis),
              data.tm.comb=data.tm.comb,
              den.fpca=den.fpca))
}

```


##cluster results
```{r,echo=FALSE, warning=FALSE}
para <- expand.grid(tractnames, c("FA","MD","Da","Dr"))
## equally spaced density 3 clusters 
misrate_FA <- sapply(1:27, function(x) m.tract.cluster(alldata.fb, x)$misrate)
misrate_MD <- sapply(28:54, function(x) m.tract.cluster(alldata.fb, x)$misrate)
misrate_Da <- sapply(55:81, function(x) m.tract.cluster(alldata.fb, x)$misrate)
misrate_Dr <- sapply(82:108, function(x) m.tract.cluster(alldata.fb, x)$misrate)

misrate_FA_q <- sapply(1:27, function(x) m.tract.cluster(alldata.fb_q, x)$misrate)
misrate_MD_q <- sapply(28:54, function(x) m.tract.cluster(alldata.fb_q, x)$misrate)
misrate_Da_q <- sapply(55:81, function(x) m.tract.cluster(alldata.fb_q, x)$misrate)
misrate_Dr_q <- sapply(82:108, function(x) m.tract.cluster(alldata.fb_q, x)$misrate)


cat("Misrate for 3 clusters:\n")
cat("No transformation: FA\n")
misrate_FA[order(misrate_FA)[1:5]]
cat("No transformation: MD\n")
misrate_MD[order(misrate_MD)[1:5]]
cat("No transformation: Da\n")
misrate_Da[order(misrate_Da)[1:5]]
cat("No transformation: Dr\n")
misrate_Dr[order(misrate_Dr)[1:5]]
cat("Corresponding tracts:\n")
cat("FA:\n")
tractnames[order(misrate_FA)[1:5]]
cat("MD:\n")
tractnames[order(misrate_MD)[1:5]]
cat("Da:\n")
tractnames[order(misrate_Da)[1:5]]
cat("Dr:\n")
tractnames[order(misrate_Dr)[1:5]]



## use quantile-density transformation corresponding to quantiles 3 clusters
misrate_FA_qdt <- sapply(1:27, function(x) m.tract.cluster(alldata.fb_qdt, x)$misrate)
misrate_MD_qdt <- sapply(28:54, function(x) m.tract.cluster(alldata.fb_qdt, x)$misrate)
misrate_Da_qdt <- sapply(55:81, function(x) m.tract.cluster(alldata.fb_qdt, x)$misrate)
misrate_Dr_qdt <- sapply(82:108, function(x) m.tract.cluster(alldata.fb_qdt, x)$misrate)

cat("Misrate for 3 clusters:\n")
cat("Quantile-density transformation: FA\n")
misrate_FA_qdt[order(misrate_FA_qdt)[1:5]]
cat("Quantile-density transformation: MD\n")
misrate_MD_qdt[order(misrate_MD_qdt)[1:5]]
cat("Quantile-density transformation: Da\n")
misrate_Da_qdt[order(misrate_Da_qdt)[1:5]]
cat("Quantile-density transformation: Dr\n")
misrate_Dr_qdt[order(misrate_Dr_qdt)[1:5]]

cat("Corresponding tracts:\n")
cat("FA:\n")
tractnames[order(misrate_FA_qdt)[1:5]]
cat("MD:\n")
tractnames[order(misrate_MD_qdt)[1:5]]
cat("Da:\n")
tractnames[order(misrate_Da_qdt)[1:5]]
cat("Dr:\n")
tractnames[order(misrate_Dr_qdt)[1:5]]

# use quantile-density transformation but using global min and max to perform standardization
misrate_FA_qdtg <- sapply(1:27, function(x) m.tract.cluster(alldata.fb_qdtg, x)$misrate)
misrate_MD_qdtg <- sapply(28:54, function(x) m.tract.cluster(alldata.fb_qdtg, x)$misrate)
misrate_Da_qdtg <- sapply(55:81, function(x) m.tract.cluster(alldata.fb_qdtg, x)$misrate)
misrate_Dr_qdtg <- sapply(82:108, function(x) m.tract.cluster(alldata.fb_qdtg, x)$misrate)


summary_res <- data.frame(Measure=c(rep("FA",5), rep("MD",5), rep("Da",5), rep("Dr",5)), 
           misrate_d=c(misrate_FA[order(misrate_FA)[1:5]], 
                           misrate_MD[order(misrate_MD)[1:5]],
                           misrate_Da[order(misrate_Da)[1:5]],
                           misrate_Dr[order(misrate_Dr)[1:5]]),
           misrate_q=c(misrate_FA_q[order(misrate_FA_q)[1:5]], 
                           misrate_MD_q[order(misrate_MD_q)[1:5]],
                           misrate_Da_q[order(misrate_Da_q)[1:5]],
                           misrate_Dr_q[order(misrate_Dr_q)[1:5]]),
           misrate_qdt=c(misrate_FA_qdt[order(misrate_FA_qdt)[1:5]], 
                           misrate_MD_qdt[order(misrate_MD_qdt)[1:5]],
                           misrate_Da_qdt[order(misrate_Da_qdt)[1:5]],
                           misrate_Dr_qdt[order(misrate_Dr_qdt)[1:5]]),
           misrate_qdtg=c(misrate_FA_qdtg[order(misrate_FA_qdtg)[1:5]], 
                           misrate_MD_qdtg[order(misrate_MD_qdtg)[1:5]],
                           misrate_Da_qdtg[order(misrate_Da_qdtg)[1:5]],
                           misrate_Dr_qdtg[order(misrate_Dr_qdtg)[1:5]]),
           tract_d=c(tractnames[order(misrate_FA)[1:5]], 
                           tractnames[order(misrate_MD)[1:5]],
                           tractnames[order(misrate_Da)[1:5]],
                           tractnames[order(misrate_Dr)[1:5]]),
           tract_q=c(tractnames[order(misrate_FA_q)[1:5]], 
                           tractnames[order(misrate_MD_q)[1:5]],
                           tractnames[order(misrate_Da_q)[1:5]],
                           tractnames[order(misrate_Dr_q)[1:5]]),
           tract_qdt=c(tractnames[order(misrate_FA_qdt)[1:5]], 
                           tractnames[order(misrate_MD_qdt)[1:5]],
                           tractnames[order(misrate_Da_qdt)[1:5]],
                           tractnames[order(misrate_Dr_qdt)[1:5]]))
```
## only cluster first two groups
```{r, echo=FALSE}
# 2 cluster density
misrate_FA2 <- sapply(1:27,  function(x) m.tract.cluster2(alldata.fb, x)$misrate)
misrate_MD2 <- sapply(28:54, function(x) m.tract.cluster2(alldata.fb, x)$misrate)
misrate_Da2 <- sapply(55:81, function(x) m.tract.cluster2(alldata.fb, x)$misrate)
misrate_Dr2 <- sapply(82:108,function(x) m.tract.cluster2(alldata.fb, x)$misrate)
cat("Misrate for 2 clusters:\n")
cat("No transformation: FA\n")
misrate_FA2[order(misrate_FA2)[1:5]]
cat("No transformation: MD\n")
misrate_MD2[order(misrate_MD2)[1:5]]
cat("No transformation: Da\n")
misrate_Da2[order(misrate_Da2)[1:5]]
cat("No transformation: Dr\n")
misrate_Dr2[order(misrate_Dr2)[1:5]]

cat("Corresponding tracts:\n")
cat("FA:\n")
tractnames[order(misrate_FA2)[1:5]]
cat("MD:\n")
tractnames[order(misrate_MD2)[1:5]]
cat("Da:\n")
tractnames[order(misrate_Da2)[1:5]]
cat("Dr:\n")
tractnames[order(misrate_Dr2)[1:5]]


#use quantile-density transformation corresponding to quantiles 2 clusters
misrate_FA_q2 <- sapply(1:27,  function(x) m.tract.cluster2(alldata.fb_q, x)$misrate)
misrate_MD_q2 <- sapply(28:54, function(x) m.tract.cluster2(alldata.fb_q, x)$misrate)
misrate_Da_q2 <- sapply(55:81, function(x) m.tract.cluster2(alldata.fb_q, x)$misrate)
misrate_Dr_q2 <- sapply(82:108,function(x) m.tract.cluster2(alldata.fb_q, x)$misrate)

misrate_FA_qdt2 <- sapply(1:27,  function(x) m.tract.cluster2(alldata.fb_qdt, x)$misrate)
misrate_MD_qdt2 <- sapply(28:54, function(x) m.tract.cluster2(alldata.fb_qdt, x)$misrate)
misrate_Da_qdt2 <- sapply(55:81, function(x) m.tract.cluster2(alldata.fb_qdt, x)$misrate)
misrate_Dr_qdt2 <- sapply(82:108,function(x) m.tract.cluster2(alldata.fb_qdt, x)$misrate)
cat("Misrate for 2 clusters:\n")
cat(" FA\n")
misrate_FA_qdt2[order(misrate_FA_qdt2)[1:5]]
cat("qd transformation: MD\n")
misrate_MD_qdt2[order(misrate_MD_qdt2)[1:5]]
cat("qd transformation: Da\n")
misrate_Da_qdt2[order(misrate_Da_qdt2)[1:5]]
cat("qd transformation: Dr\n")
misrate_Dr_qdt2[order(misrate_Dr_qdt2)[1:5]]

cat("Corresponding tracts:\n")
cat("Quantile-density transformation: FA\n")
tractnames[order(misrate_FA_qdt2)[1:5]]
cat("Quantile-density transformation: MD\n")
tractnames[order(misrate_MD_qdt2)[1:5]]
cat("Quantile-density transformation: Da\n")
tractnames[order(misrate_Da_qdt2)[1:5]]
cat("Quantile-density transformation: Dr\n")
tractnames[order(misrate_Dr_qdt2)[1:5]]

#use quantile-density transformation corresponding to quantiles 2 clusters but with global min and max 
misrate_FA_qdtg2 <- sapply(1:27,  function(x) m.tract.cluster2(alldata.fb_qdtg, x)$misrate)
misrate_MD_qdtg2 <- sapply(28:54, function(x) m.tract.cluster2(alldata.fb_qdtg, x)$misrate)
misrate_Da_qdtg2 <- sapply(55:81, function(x) m.tract.cluster2(alldata.fb_qdtg, x)$misrate)
misrate_Dr_qdtg2 <- sapply(82:108,function(x) m.tract.cluster2(alldata.fb_qdtg, x)$misrate)



summary_res2 <- data.frame(Measure=c(rep("FA",5), rep("MD",5), rep("Da",5), rep("Dr",5)), 
           misrate_d=c(misrate_FA2[order(misrate_FA2)[1:5]], 
                           misrate_MD2[order(misrate_MD2)[1:5]],
                           misrate_Da2[order(misrate_Da2)[1:5]],
                           misrate_Dr2[order(misrate_Dr2)[1:5]]),
           misrate_q=c(misrate_FA_q2[order(misrate_FA_q2)[1:5]], 
                           misrate_MD_q2[order(misrate_MD_q2)[1:5]],
                           misrate_Da_q2[order(misrate_Da_q2)[1:5]],
                           misrate_Dr_q2[order(misrate_Dr_q2)[1:5]]),
           misrate_qdt=c(misrate_FA_qdt2[order(misrate_FA_qdt2)[1:5]], 
                           misrate_MD_qdt2[order(misrate_MD_qdt2)[1:5]],
                           misrate_Da_qdt2[order(misrate_Da_qdt2)[1:5]],
                           misrate_Dr_qdt2[order(misrate_Dr_qdt2)[1:5]]),
           misrate_qdtg=c(misrate_FA_qdtg2[order(misrate_FA_qdtg2)[1:5]], 
                           misrate_MD_qdtg2[order(misrate_MD_qdtg2)[1:5]],
                           misrate_Da_qdtg2[order(misrate_Da_qdtg2)[1:5]],
                           misrate_Dr_qdtg2[order(misrate_Dr_qdtg2)[1:5]]),
           tract_d=c(tractnames[order(misrate_FA2)[1:5]], 
                           tractnames[order(misrate_MD2)[1:5]],
                           tractnames[order(misrate_Da2)[1:5]],
                           tractnames[order(misrate_Dr2)[1:5]]),
           tract_q=c(tractnames[order(misrate_FA_q2)[1:5]], 
                           tractnames[order(misrate_MD_q2)[1:5]],
                           tractnames[order(misrate_Da_q2)[1:5]],
                           tractnames[order(misrate_Dr_q2)[1:5]]),
           tract_qdt=c(tractnames[order(misrate_FA_qdt2)[1:5]], 
                           tractnames[order(misrate_MD_qdt2)[1:5]],
                           tractnames[order(misrate_Da_qdt2)[1:5]],
                           tractnames[order(misrate_Dr_qdt2)[1:5]]),
           tract_qdtg=c(tractnames[order(misrate_FA_qdtg2)[1:5]], 
                           tractnames[order(misrate_MD_qdtg2)[1:5]],
                           tractnames[order(misrate_Da_qdtg2)[1:5]],
                           tractnames[order(misrate_Dr_qdtg2)[1:5]]))

```
## Visualized variation modes
```{r}
#fb_score <- lapply(1:dim(para)[1], function(x) m.tract.cluster2(alldata.fb, x)$data.tm.comb)
#fb_score <- do.call(rbind, fb_score) #9280x209
#fb_scorefpca <- lapply(1:dim(para)[1], function(x) m.tract.cluster2(alldata.fb, x)$den.fpca)

var_mode <- function(data, m, t, title, qd){
loc_m <- which(para[,2]==m)
loc <- loc_m[which(para[loc_m,1]==t)] #loc in the para that is tract=='unc_r', measure=='Da'
fb_scorefpca <- lapply(loc, function(x) m.tract.cluster2(data, x)$den.fpca)

eigenf <- fb_scorefpca[[1]]$efunctions*sqrt(200)
eigens <- fb_scorefpca[[1]]$scores/sqrt(200)
mean <- fb_scorefpca[[1]]$mu
pc1_score_q <- quantile(eigens[,1], c(0.1, 0.25, 0.75, 0.95))
modes1 <- cbind(t(as.vector(pc1_score_q)%*% t(eigenf[,1])) + mean, mean) #200x5
pc2_score_q <- quantile(eigens[,2], c(0.1, 0.25, 0.75, 0.95))
modes2 <- cbind(t(as.vector(pc2_score_q)%*% t(eigenf[,2])) + mean, mean) #200x5
pc3_score_q <- quantile(eigens[,3], c(0.1, 0.25, 0.75, 0.95))
modes3 <- cbind(t(as.vector(pc3_score_q)%*% t(eigenf[,3])) + mean, mean) #200x5

tran_back <- function(den.tran.y, den.quantiles){
  thetax <- sum((head(exp(den.tran.y), -1)+tail(exp(den.tran.y), -1))/2*(diff(den.quantiles/100)[1]))

  tranback_x <- sapply(1:200, function(x){
    sum((head(exp(den.tran.y)[1:x], -1)+tail(exp(den.tran.y)[1:x],-1))/2*(diff(den.quantiles/100)[1]))/thetax
  })
  tranback_y <- exp(-den.tran.y)*thetax
  return(list(tranback_x=tranback_x,
              tranback_y=tranback_y,
              thetax=thetax)) 
}

if(qd==0){
matplot(den.quantiles/100, modes1, main=paste0("1st PC modes, ",title, "\n tract:", t, ", measure:", m),
        xlab="t", ylab=expression(psi(t)), 
        col=c("red","pink","green","blue", "black"), type="l", 
        lty=c(1,1,1,1,1), lwd=c(1,1,1,1,1.5))
abline(h=0, col="grey", lwd=1, lty=2)
legend("topleft", c(expression(alpha[1]), expression(alpha[2]), expression(alpha[3]), expression(alpha[4]), expression(alpha[5])), col=c("red","pink","black","green","blue"), lty=c(1,1,1,1,1), lwd=c(1,1,1.5,1,1), cex=0.8, bty='n')

matplot(den.quantiles/100, modes2, main=paste0("2nd PC modes, ",title, "\n tract:", t, ", measure:", m),
        xlab="t", ylab=expression(psi(t)), 
        col=c("red","pink","green","blue", "black"), type="l",
        lty=c(1,1,1,1,1), lwd=c(1,1,1,1,1.5))
abline(h=0, col="grey", lwd=1, lty=2)
legend("topleft", c(expression(alpha[1]), expression(alpha[2]), expression(alpha[3]), expression(alpha[4]), expression(alpha[5])), col=c("red","pink","black","green","blue"), lty=c(1,1,1,1,1), lwd=c(1,1,1.5,1,1), cex=0.8, bty='n')

}
if(qd==1){
tran_den <- lapply(data.frame(modes1), function(x){tran_back(x, den.quantiles)}) 
plot(tran_den[[1]]$tranback_x, tran_den[[1]]$tranback_y, type="l",main=paste0("1st PC modes, ",title, "\n tract:", t, ", measure:", m), col="red", ylim=c(0, max(c(tran_den[[1]]$tranback_y, tran_den[[2]]$tranback_y,
                                tran_den[[3]]$tranback_y, tran_den[[4]]$tranback_y,
                                tran_den[[5]]$tranback_y))),
     xlab="x", ylab=expression(psi^{-1}(x)))
abline(h=0, col="grey", lwd=1, lty=2)
lines(tran_den[[2]]$tranback_x, tran_den[[2]]$tranback_y, type="l", col="pink")
lines(tran_den[[3]]$tranback_x, tran_den[[3]]$tranback_y, type="l", col="green")
lines(tran_den[[4]]$tranback_x, tran_den[[4]]$tranback_y, type="l", col="blue")
lines(tran_den[[5]]$tranback_x, tran_den[[5]]$tranback_y,col="black", lwd=1.5)
legend("topleft", c(expression(alpha[1]), expression(alpha[2]), expression(alpha[3]), expression(alpha[4]), expression(alpha[5])), col=c("red","pink","black","green","blue"), lty=c(1,1,1,1,1), lwd=c(1,1,1.5,1,1), cex=0.8, bty='n')

tran_den2 <- lapply(data.frame(modes2), function(x){tran_back(x, den.quantiles)}) 
plot(tran_den2[[1]]$tranback_x, tran_den2[[1]]$tranback_y, type="l",main=paste0("2nd PC modes, ",title, "\n tract:", t, ", measure:", m),
     xlab="x", ylab=expression(psi^{-1}(x)), 
     col="red", ylim=c(0, max(c(tran_den2[[1]]$tranback_y, tran_den2[[2]]$tranback_y, 
                                tran_den2[[3]]$tranback_y, tran_den2[[4]]$tranback_y,
                                tran_den2[[5]]$tranback_y))))
abline(h=0, col="grey", lwd=1, lty=2)
lines(tran_den2[[2]]$tranback_x, tran_den2[[2]]$tranback_y, type="l", col="pink")
lines(tran_den2[[3]]$tranback_x, tran_den2[[3]]$tranback_y, type="l", col="green")
lines(tran_den2[[4]]$tranback_x, tran_den2[[4]]$tranback_y, type="l", col="blue")
lines(tran_den2[[5]]$tranback_x, tran_den2[[5]]$tranback_y, col="black", lwd=1.5)
legend("topleft", c(expression(alpha[1]), expression(alpha[2]), expression(alpha[3]), expression(alpha[4]), expression(alpha[5])), col=c("red","pink","black","green","blue"), lty=c(1,1,1,1,1), lwd=c(1,1,1.5,1,1), cex=0.8, bty='n')

tran_den3 <- lapply(data.frame(modes3), function(x){tran_back(x, den.quantiles)}) 
plot(tran_den3[[1]]$tranback_x, tran_den3[[1]]$tranback_y, type="l",main=paste0("3rd PC modes, ",title, "\n tract:", t, ", measure:", m),
     xlab="x", ylab=expression(psi^{-1}(x)), 
     col="red", ylim=c(0, max(c(tran_den3[[1]]$tranback_y, tran_den3[[2]]$tranback_y, 
                                tran_den3[[3]]$tranback_y, tran_den3[[4]]$tranback_y,
                                tran_den3[[5]]$tranback_y))))
abline(h=0, col="grey", lwd=1, lty=2)
lines(tran_den3[[2]]$tranback_x, tran_den3[[2]]$tranback_y, type="l", col="pink")
lines(tran_den3[[3]]$tranback_x, tran_den3[[3]]$tranback_y, type="l", col="green")
lines(tran_den3[[4]]$tranback_x, tran_den3[[4]]$tranback_y, type="l", col="blue")
lines(tran_den3[[5]]$tranback_x, tran_den3[[5]]$tranback_y, col="black", lwd=1.5)
legend("topleft", c(expression(alpha[1]), expression(alpha[2]), expression(alpha[3]), expression(alpha[4]), expression(alpha[5])), col=c("red","pink","black","green","blue"), lty=c(1,1,1,1,1), lwd=c(1,1,1.5,1,1), cex=0.8, bty='n')

}
}

locM <- which.min(c(min(misrate_FA_qdt2), min(misrate_MD_qdt2), min(misrate_Da_qdt2), min(misrate_Dr_qdt2)))
minM <- c('FA', 'MD', 'Da', 'Dr')[locM]
minT <- c(as.character(tractnames)[order(misrate_FA_qdt2)[1]], 
          as.character(tractnames)[order(misrate_MD_qdt2)[1]],
          as.character(tractnames)[order(misrate_Da_qdt2)[1]],
          as.character(tractnames)[order(misrate_Dr_qdt2)[1]])[locM]

var_mode(alldata.fb, m=minM, t=minT, title='Ordinary FPCA', qd=0)

var_mode(alldata.fb_qdt, m=minM, t=minT, title='log quantile-density transformation', qd=1)

var_mode(alldata.fb_qdtg, m="FA", t="cst_r", title='log quantile-density transformation', qd=1) # this plot is based on the "alldata_qdtg4.Rdata"



## add the density plot that what the true label looks like 
den_select2 <- alldata.fb_qdt %>% filter(measure=="Dr", tract=="ar_l", group!=3) %>% select(den1:den200,group)
group2_<- den_select2$group
den_select2 <- select(den_select2, -group)
matplot(den.quantiles/100, t(den_select2), type="l", lty=1, col=ifelse(group2_==1, "red","green"))


den_select2 <- alldata.fb %>% filter(measure=="Dr", tract=="ar_l", group!=3, player %in% footballname) %>% select(den1:den200,group)
group2_<- den_select2$group
den_select2 <- select(den_select2, -group)
matplot(den.quantiles/100, t(den_select2), type="l", lty=1, col=ifelse(group2_==1, "red","green"), xlab="x", ylab="f(x)", main="True density function:\n tract:ar_l, measure:Dr")

den_select2 <- alldata.fb %>% filter(measure=="FA", tract=="cst_l", group!=3) %>% select(den1:den200,group)
group2_<- den_select2$group
den_select2 <- select(den_select2, -group)
matplot(den.quantiles/100, t(den_select2), type="l", lty=1, col=ifelse(group2_==1, "red","green"), xlab="x", ylab="f(x)", main="True density function:\n tract:cst_l, measure:FA")
 legend("topright", c("group1", "group2"), col=c("red", "green"), lty=c(1,1))
 
t <- "cst_l"
  m <- "FA"
  data <- alldata.fb_qdtg
  data.tm <- data %>% filter(tract==t, measure==m, group != 3) #86x204
  den.tm <- select(data.tm, den1:den200)    #86x200
  #den.tm <- den.tm - as.vector(rep(1, dim(den.tm)[1])) %*% t(colMeans(den.tm))
  den.fpca <- refund::fpca.face(data.matrix(den.tm), center = TRUE, argvals = (1:200)/200, knots = 7, pve = 0.95, p = 5, lambda = 0)
  den.score <- den.fpca$scores/sqrt(dim(den.tm)[2]) #73x8
  
  den.cluster <- kmeans(den.score[,1:2], centers = 2, iter.max = 100)
    permute <- function(vec, p1, p2){
      for(i in 1:length(vec)){
        if(vec[i]==1){vec[i] <- p1}else{vec[i] <- p2}
      }
      return(vec)
    }
    perm1 <- permute(den.cluster$cluster, 1,2)
    perm2 <- permute(den.cluster$cluster, 2,1)
    perm <- cbind(perm1, perm2)
    misrate <- c(sum(perm1!=data.tm$group)/length(data.tm$group),
                   sum(perm2!=data.tm$group)/length(data.tm$group))
    
    group.tm <- perm[,order(misrate)[1]]
    matplot(den.quantiles/100, t(den_select2), type="l", lty=1, col=ifelse(group.tm==1, "red","green"), xlab="x", ylab="f(x)", main=paste0("Clustered density function:\n tract:cst_l, measure:FA(misrate:", round(min(misrate),digits=2),")"))
    legend("topright", c("group1", "group2"), col=c("red", "green"), lty=c(1,1))


plot(den.score[,1], den.score[,2], col=ifelse(group2_==1, "red", "green"), 
     xlab="PC1", ylab="PC2",
     main=paste0("tract:cst_l, measure:FA(misrate:", round(min(misrate),digits=2),")"))
points(den.score[,1], den.score[,2], pch=3, col=ifelse(group.tm==1, "red", "green"))    
    
 
plot(den.score[,1], den.score[,2], col=ifelse(group2_==1, "red", "green"), 
     xlab="PC1", ylab="PC2",
     main=paste0("tract:cst_l, measure:FA(misrate:", round(min(misrate),digits=2),")"))
points(den.score[,1], den.score[,2], pch=3, col=ifelse(a=="group1", "red", "green"))    
    
    
summary_arlDr <- sapply(footballloc, function(x){
  sub.index <- x 
  subi_obs <- fread(fulldir[x])
  tract.j <- subi_obs %>% filter(Tract=="ar_l") %>% select(-Tract, "Dr")
  c(min(tract.j), max(tract.j))
})
min(summary_arlDr[1,])
max(summary_arlDr[2,])

persub_arlDr <- sapply(1:length(fulldir), function(x){
  #sub.index: 1~195
  #measure:any measurement subset from c("FA", "MD", "Da", "Dr")
  sub.index <- x 
  sub_dir <- fulldir[sub.index]
  sub.i <- read.csv(sub_dir)
  sub.i.measure <- sub.i %>% select(Tract,"Dr") %>% filter(Tract=="ar_l")
  # den.est is a list of length(tracts in sub.index), each element is a matrix of length(measure)x203
  norm_subi <- (sub.i.measure$Dr - min(summary_arlDr[1,]))/(max(summary_arlDr[2,]) - min(summary_arlDr[1,]))
  den.col <- density(norm_subi, n=length(norm_subi), na.rm=TRUE)
  interpl <- approx(den.col$x, den.col$y, den.quantiles/100)
  interpl$y
}) 


matplot(den.quantiles/100, persub_arlDr[, arlDr_group!=3], type="l", lty=1, col=ifelse(arlDr_group[arlDr_group!=3]==1, "red","green"), xlab="x", ylab="f(x)", main="True density function:\n tract:ar_l, measure:Dr")






```
